Code examples to accompany "If the Null Fits, You Must Omit

### 2025-11-19 GroupPolarizationStatmod/paper/figures.tex
Created paper/code_examples.qmd and continued on method

>>>>

\begin{figure}
  \centering
    % \includegraphics[width=0.8\textwidth]{Figures/Model/data-model-flow.png}
  \includegraphics[width=\textwidth]{Figures/Model/SoftwareMap_Mashup.png}
  \caption{\textbf{Simulation and analysis pipeline}
  for finding when measurement artifacts.  
  Arrows show the flow of input and output data from parameter estimation.
  First, we used the web app (A) to find null-polarization parameters for 
  each experimental condition, $\nu_e$,
  that ordinal measurements distort to look like polarization
  (B—described in Algorithm~\ref{alg:hillclimbing}).
  We fit ordered probit models to simulated null-polarization outcomes,
  seeded by the $\nu$ found in the previous step (C). Analysis and 
  visualization routines finish the pipeline (D).}
  \label{fig:data-model-flow}
\end{figure}

### 2025-11-19 GroupPolarizationStatmod/paper/method.tex
Created paper/code_examples.qmd and continued on method

The null-polarization distributions are represented by

@@

\subsubsection{Simulated opinion measurements}

@@

All bins have a width of 1 in latent opinion space, except for the two most
extreme bins at opposite ends of the sentiment spectrum—this works because
bin values are always separated by 1. All continuous latent opinion values
$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
value $b_1$—so, $b_1$'s lower threshold is $\theta_0 = -\infty$.  Similarly,
all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$.  We
can define all thresholds compactly as \cite{Liddell2018}
\begin{equation}
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_1 + \frac{j + 1}{2} & \text{ otherwise.}
\end{cases}
\label{eq:thresholds}
\end{equation}
\noindent

@@

\begin{center}
\vspace{1.5em}
[ FIGURE \ref{fig:distros} ABOUT HERE ]
\vspace{1.5em}
\end{center}

@@

\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}

>>>>

The null-polarization distributions are represented by $\nu$.

@@

\subsubsection{Simulated group polarization measurements}

@@

All bins have a width of 1 in latent opinion space, except for the two most
extreme bins at opposite ends of the sentiment spectrum—this works because
bin values are always separated by 1. All continuous latent opinion values
$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
value $b_1$—so, $b_1$'s lower threshold is $\theta_0 = -\infty$.  Similarly,
all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$.  We
can define all thresholds compactly as \cite{Liddell2018}
\begin{equation}
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
\label{eq:thresholds}
\end{equation}
\noindent

@@

@@

\subsubsection{Simulated spurious group polarization}

Group polarization measures radicalization in experimental echo chambers in terms of
the difference between mean group opinions before and after the group discusses some
survey topic. The following model shows the problem of how the measurement scheme 
above can make \emph{mere agreement} spuriously look like polarization.

To make things more exact, we define a \emph{null model}, $\nu$, of mere agreement
defined as a 3-tuple: a constant mean for the pre- and post-discussion latent opinion
distributions, and a pre-deliberation latent opinion variance that is greater than the
post-deliberation variance to represent finding agreement,
\begin{equation}
  \nu = (\mu,~\sigmapre,~\sigmapost).
  \label{eq:nu-definition}
\end{equation}
\noindent
Group polarization is \emph{spurious} in this case if measurement clipping causes 
the observed difference in pre- and post-discussion opinions to be greater than zero,
i.e., $g > 0$ (Equation \ref{eq:group-polarization}, 
even if the true latent change is zero.

We will show how this arrangement generates spurious polarization in an
experimental condition in our corpus, 
the \emph{Americans} experimental condition in
\textcite{Moscovici1969}, to which we gave the Experiment ID
\texttt{Moscovici1969\_Americans}. In that condition, the authors used a seven-bin
scale ($B=7$) with numerical values from -3 to +3 indicating opinions from ``strongly
disagree'' to ``strongly agree''. Pre-consensus opinions were reported to have
mean -0.6 and the post-consensus mean was -1.04 \cite[p. 132, Table 4]{Moscovici1969}.

And now we will show that a null model with
$\nu = (\mu = -1.2, \sigmapre = 4.3, and \sigmapost = 1.9)$ generates spurious
polarization with exactly the pre- and post-discussion means Moscovici and Zavalloni
report (Figure~\ref{fig:distros}). First, calculate the vector of probability density
values for each bin, $f[b]$, for the pre- and post-discussion latent opinion
distributions using Equation \ref{eq:} (calculated with
[`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R\#L43) from 
the [project code repository](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod):
\begin{subequations}
  \begin{equation}
    f_\mathrm{pre}[b] = 
  \end{equation}
  \begin{equation}
    f_\mathrm{post}[b] = 
  \end{equation}
\end{subequations}

\begin{center}
\vspace{1.5em}
[ FIGURE \ref{fig:distros} ABOUT HERE ]
\vspace{1.5em}
\end{center}
\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}
